---
title: "Case Study 2-Employees"
author: "Xavier Mojica"
date: "2022-11-17"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Employee Data
## Presenting Findings on Employees 
### R Data Manipuilationn and Import. Pre load packages.
```{r}
library(ggplot2)
library(stringi)
library(readxl)
library(Metrics)
library(XML)
library(rvest)
library(naniar)
library(e1071)
library(caret)
library(class)
library(tidyverse)
library(stringr)
library(dplyr)
library(GGally)
library(tidyr)
library(knitr)
library(rmarkdown)
library(psych)
library(fastDummies)
```
#Reading in File
```{r}
#Reading our Beer and Brewery Files
employeeData = read.csv('CaseStudy2-data (1).csv', header = TRUE)

#Checking data
head(employeeData)
```

#Cleaning Data
```{r}
##Cleaning up data by removing NAs. 
is.na(employeeData)
CleanEmployeeData = na.omit(employeeData)

#summarizing data
str(CleanEmployeeData)
summary(CleanEmployeeData)
```

#Changing OverTime to 1s and 0s
```{r}
CleanEmployeeData$OverTimeScale = ifelse(CleanEmployeeData$OverTime =="Yes", 1,0) 
```

#Scaling Columns
```{r}
CleanEmployeeData$MonthlyIncomeScale = scale(CleanEmployeeData$MonthlyIncome)
CleanEmployeeData$AgeScale = scale(CleanEmployeeData$Age)
CleanEmployeeData$EnvironmentSatisfactionScale = scale(CleanEmployeeData$EnvironmentSatisfaction)
CleanEmployeeData$OverTimeScale = scale(CleanEmployeeData$OverTimeScale)

CleanEmployeeData$YearsAtCompanyScale = scale(CleanEmployeeData$YearsAtCompany)

head(CleanEmployeeData)

```




#Ploting
```{r,echo=FALSE, message=FALSE,warning=FALSE,results="hide"}

cor(CleanEmployeeData[,unlist(lapply(CleanEmployeeData, is.numeric))])

#more refined by removing unnecessary columns
#cor(CleanEmployeeData[,c(2,5,7,8,12,14,15,16,18, 20,21,25,26,27,29,30, 31,32,33,34,35,36,37)])
pairs.panels(CleanEmployeeData[,c(2,5,7,8,12,14,15,16,18, 20,21,25,26,27,29,30,31,32,33,34,35,36,37)])

#ploting correlation 
#pairs.panels(CleanEmployeeData[,c(2,5,7,8,12,14,15,16,18)])
#pairs.panels(CleanEmployeeData[,c(20,21,25,26,27,29,30,31,32,33,34,35,36,37)])
#can use a facet ggplot with like 5 variables i.e. , r shinny allows or packages in R that does the same with custom tables , display on subset of correlations 
ggpairs(CleanEmployeeData[,c(3,2,12,20,18)], aes(col = Attrition))

ggpairs(CleanEmployeeData[,c(3,25,27,35)], aes(col = Attrition))

ggpairs(CleanEmployeeData[,c(3,24,16,33 )], aes(col = Attrition))
```

#Correlation Plots
```{r}

#Attrition Count
CleanEmployeeData %>% ggplot(aes(x=Attrition,fill=Attrition)) + labs(title = "Attrition Count", x = "Attrition") + geom_bar()

#Attrition V. Monthly Income
CleanEmployeeData %>% ggplot(aes(x=MonthlyIncome,fill=Attrition)) + ggtitle("Attrition V. Monthly Income") + geom_histogram()

#Attrition V. Monthly Rate
CleanEmployeeData %>% ggplot(aes(x=MonthlyRate,fill=Attrition)) + geom_histogram() + ggtitle("Attrition Vs. Monthly Rate") 

#Attrition V. OverTime
CleanEmployeeData %>% ggplot(aes(x=OverTime,fill=Attrition)) + geom_bar(position="fill") + ggtitle("Attrition Vs. Overtime")


# Attrition V. Job Satisfaction
CleanEmployeeData %>% ggplot(aes(x=JobSatisfaction,fill=Attrition))+ geom_bar() + labs(title = "Attrition Vs. Job Satisfaction")

#Attrition V. Age
CleanEmployeeData %>% ggplot(aes(x=Age,fill=Attrition)) + ggtitle("Attrition Vs. Age") + geom_histogram()

# Attrition V. Relationship Satisfaction
CleanEmployeeData %>% ggplot(aes(x=RelationshipSatisfaction,fill=Attrition)) + geom_histogram() + ggtitle("Attrition Vs. RelationshipSatisfaction")

# Attrition V. Work Life BalanceJob
CleanEmployeeData %>%ggplot(aes(x=WorkLifeBalance,fill=Attrition)) + geom_histogram() + labs( title = "Attrition Vs. WorkLifeBalance") 

# By Job Role
ggplot(CleanEmployeeData, aes(x = JobRole, fill = Attrition)) + geom_bar(position = "fill") + theme(axis.text.x = element_text(angle = 60, vjust = 1, hjust= 1))

# Total of Employees By Job Role 
ggplot(CleanEmployeeData, aes(x = JobRole, fill = JobRole)) + geom_bar() + theme(axis.text.x = element_text(angle = 60, vjust = 1, hjust=1)) + labs(title = "Employees by Job Role", fill = "Job Role")

#HourlyRate v. Job Role 
ggplot(CleanEmployeeData, aes(x = JobRole, y = HourlyRate, fill = factor(JobRole))) + theme(axis.text.x = element_text(angle = 60, vjust = 1, hjust=1)) + labs(title = "Hourly Rate of Employees by Job Role", fill = "Job Roles") + geom_boxplot()

#MonthlyIncome v. JobRole
ggplot(CleanEmployeeData, aes(x = JobRole, y = MonthlyIncome, fill = factor(JobRole))) + theme(axis.text.x = element_text(angle = 60, vjust = 1, hjust=1)) + labs(title = "Montly Income of Employees by Job Role", fill = "Job Roles") + geom_boxplot()

#Job Role v. Attrition 
ggplot(CleanEmployeeData, aes(x = JobRole, fill = Attrition)) + theme(axis.text.x = element_text(angle = 60, vjust = 1, hjust=1)) + labs(title = "Attrition by Job Role", fill = "Attrition") + geom_bar(position = "fill")

#By Job Level 
ggplot(CleanEmployeeData, aes(x = JobLevel, fill = Attrition)) + geom_bar(position = "fill") + theme(axis.text.x = element_text(angle = 60, vjust = 1, hjust=1))

```


#Testing to find K 
```{r}
# Loop for many k and the average of many training / test partition

set.seed(4)

iterations = 500
numks = 30
masterAcc = matrix(nrow = iterations, ncol = numks)

for(j in 1:iterations)
{
  accs = data.frame(accuracy = numeric(30), k = numeric(30))
  trainIndices = sample(1:dim(CleanEmployeeData)[1], round(0.70* dim(CleanEmployeeData)[1]))
  train = CleanEmployeeData[trainIndices,]
  test = CleanEmployeeData[-trainIndices,]
  for(i in 1:numks)
  {
    classifications = knn(train[,c(37,38,40)], test[,c(37,38,40)], train$Attrition, prob = TRUE, k = i)   #37,38,40
    table(classifications, test$Attrition)
    cm = confusionMatrix(table(classifications, test$Attrition))
    masterAcc[j,i] = cm$overall[1]
  }
}
MeanAcc = colMeans(masterAcc)
plot(seq(1,numks,1),MeanAcc, type = "l")

```
#Creating the KNN Model
```{r}
#KNN
classifications = knn(train[,c(37,38,40)], test[,c(37,38,40)], train$Attrition, prob = TRUE, k =15) #9 previously 
classifications

table(classifications, test$Attrition)
cm = confusionMatrix(table(classifications, test$Attrition))
cm
# Accuracy : 0.8621, Sensitivity : 0.9381, Specificity : 0.3714  
```



#Naive Bayes
```{r}

model = CleanEmployeeData %>% select(c("JobRole","MonthlyIncome","OverTime","EnvironmentSatisfaction","Age", "DailyRate","JobLevel","JobInvolvement","TotalWorkingYears", "Attrition"))

trainIndices = sample(1:dim(model)[1],round(.70 * dim(model)[1]))
train = model[trainIndices,]
test = model[-trainIndices,]

set.seed(7)
classifier1 = naiveBayes(model[,c(2,3,7)],model$Attrition)
pred = predict(classifier1,newdata=test)
CM = confusionMatrix(table(test$Attrition,pred))


CM
#Accuracy 87.74%,  Sensitivity : 0.9068  , Specificity : 0.6000  
```


#Finding Correlations  
```{r}
cor(CleanEmployeeData$MonthlyIncome ,CleanEmployeeData[,c(2,5,7,8,12,14,15,16,18,20,21,25,26,27,29,30, 31,32,33,34,35,36,37)])

```

#Preparing with Dummy Variables 
```{r}
CleanEmployeeData1 = CleanEmployeeData
CleanEmployeeData1 = dummy_cols(CleanEmployeeData, select_columns = c("BusinessTravel", "Department", "EducationField", "Gender", "JobRole"))

CleanEmployeeData1$BusinessTravelTF = ifelse(CleanEmployeeData1$`BusinessTravel_Non-Travel` == 1, 0,1)

CleanEmployeeData1$JobRoleTF = ifelse(CleanEmployeeData1$JobRole_Manager == 1|CleanEmployeeData1$`JobRole_Sales Executive` == 1|CleanEmployeeData1$`JobRole_Research Director` == 1,0,1)

CleanEmployeeData1[, c(39:66)] = CleanEmployeeData1[,c(39:66)] %>% mutate_if(sapply(CleanEmployeeData1[, c(39:66)], is.numeric), as.factor)

lm_SalaryDF = CleanEmployeeData1[,c(2,5,7,8,12,14:16,18,20,21,25:27, 29:37, 39:64)]

lmSalaryMod = lm(MonthlyIncome~., lm_SalaryDF)

summary(lmSalaryMod)


```



#Reading in  No Slary Data
```{r}
EmployeeDataSalary = read_excel('CaseStudy2CompSet No Salary.xlsx')
```

#Cleaning No Slary Data
```{r}
CleanEmployeeDataSalary = na.omit(EmployeeDataSalary)

head(CleanEmployeeDataSalary)

```

```{r}
cor(CleanEmployeeData$MonthlyIncome ,CleanEmployeeData[,c(2,5,7,8,12,14,15,16,18, 20,21,25,26,27,29,30, 31,32,33,34,35,36,37)])

#cor(CleanEmployeeDataSalary[,unlist(lapply(CleanEmployeeDataSalary, is.numeric))])

```


#RMSE
```{r}
#str(CleanEmployeeData1)
#c(16,20,21,26,28,35,39,41,44,56,57,59
#c(16,20,21,26,30,35,42,44, 49,61, 62, 63)
#(16,30,59,61,42,26,35,63)

set.seed(7)

lm_SalaryDF = CleanEmployeeData1[,c(16,20,30,59,61,42,26,35,63)] 

lmSalaryMod = lm(MonthlyIncome~., lm_SalaryDF)

summary(lmSalaryMod)



```







 $ ID                      : int  1 2 3 4 5 6 7 8 9 10 ...
 $ Age                     : int  32 40 35 32 24 27 41 37 34 34 ...
 $ Attrition               : chr  "No" "No" "No" "No" ...
 $ BusinessTravel          : chr  "Travel_Rarely" "Travel_Rarely" "Travel_Frequently" "Travel_Rarely" ...
 $ DailyRate               : int  117 1308 200 801 567 294 1283 309 1333 653 ...
 $ Department              : chr  "Sales" "Research & Development" "Research & Development" "Sales" ...
 $ DistanceFromHome        : int  13 14 18 1 2 10 5 10 10 10 ...
 $ Education               : int  4 3 2 4 1 2 5 4 4 4 ...
 $ EducationField          : chr  "Life Sciences" "Medical" "Life Sciences" "Marketing" ...
 $ EmployeeCount           : int  1 1 1 1 1 1 1 1 1 1 ...
 $ EmployeeNumber          : int  859 1128 1412 2016 1646 733 1448 1105 1055 1597 ...
 $ EnvironmentSatisfaction : int  2 3 3 3 1 4 2 4 3 4 ...
 $ Gender                  : chr  "Male" "Male" "Male" "Female" ...
 $ HourlyRate              : int  73 44 60 48 32 32 90 88 87 92 ...
 $ JobInvolvement          : int  3 2 3 3 3 3 4 2 3 2 ...
 $ JobLevel                : int  2 5 3 3 1 3 1 2 1 2 ...
 $ JobRole                 : chr  "Sales Executive" "Research Director" "Manufacturing Director" "Sales Executive" ...
 $ JobSatisfaction         : int  4 3 4 4 4 1 3 4 3 3 ...
 $ MaritalStatus           : chr  "Divorced" "Single" "Single" "Married" ...
 $ MonthlyIncome           : int  4403 19626 9362 10422 3760 8793 2127 6694 2220 5063 ...
 $ MonthlyRate             : int  9250 17544 19944 24032 17218 4809 5561 24223 18410 15332 ...
 $ NumCompaniesWorked      : int  2 1 2 1 1 1 2 2 1 1 ...
 $ Over18                  : chr  "Y" "Y" "Y" "Y" ...
 $ OverTime                : chr  "No" "No" "No" "No" ...
 $ PercentSalaryHike       : int  11 14 11 19 13 21 12 14 19 14 ...
 $ PerformanceRating       : int  3 3 3 3 3 4 3 3 3 3 ...
 $ RelationshipSatisfaction: int  3 1 3 3 3 3 1 3 4 2 ...
 $ StandardHours           : int  80 80 80 80 80 80 80 80 80 80 ...
 $ StockOptionLevel        : int  1 0 0 2 0 2 0 3 1 1 ...
 $ TotalWorkingYears       : int  8 21 10 14 6 9 7 8 1 8 ...
 $ TrainingTimesLastYear   : int  3 2 2 3 2 4 5 5 2 3 ...
 $ WorkLifeBalance         : int  2 4 3 3 3 2 2 3 3 2 ...
 $ YearsAtCompany          : int  5 20 2 14 6 9 4 1 1 8 ...
 $ YearsInCurrentRole      : int  2 7 2 10 3 7 2 0 1 2 ...
 $ YearsSinceLastPromotion : int  0 4 2 5 1 1 0 0 0 7 ...
 $ YearsWithCurrManager    : int  3 9 2 7 3 7 3 0 0 7 ...

#No Salary Data
```{r}
CleanEmployeeDataSalary1 = CleanEmployeeDataSalary


CleanEmployeeDataSalary1 = dummy_cols(CleanEmployeeData, select_columns = c("BusinessTravel", "Department", "EducationField", "Gender", "JobRole"))

CleanEmployeeDataSalary1$BusinessTravelTF = ifelse(CleanEmployeeDataSalary1$`BusinessTravel_Non-Travel` == 1, 0,1)

CleanEmployeeDataSalary1$JobRoleTF = ifelse(CleanEmployeeDataSalary1$JobRole_Manager == 1|CleanEmployeeDataSalary1$`JobRole_Sales Executive` == 1|CleanEmployeeDataSalary1$`JobRole_Research Director` == 1,0,1)

CleanEmployeeDataSalary1[, c(39:66)] = CleanEmployeeDataSalary1[,c(39:62)] %>% mutate_if(sapply(CleanEmployeeDataSalary1[, c(39:62)], is.numeric), as.factor)

lm_SalaryDF = CleanEmployeeDataSalary1[,c(2,5,7,8,12,14:16,18,20,21,25:27, 29:37, 39:62)]

lmSalaryMod = lm(MonthlyIncome~., lm_SalaryDF)

summary(lmSalaryMod)


```


#Predicting Salary
```{r}

SalaryPredict= predict(lmSalaryMod,newdata = CleanEmployeeDataSalary1)
SalaryPredict
```

#Saving Predicted Salary to File 
```{r}
CleanEmployeeDataSalary1$PredictedSalary = (SalaryPredict)
PredictionSalaryDF = data.frame(c)

write.csv(PredictionSalaryDF, "SalaryPrediction.cvs", row.names = TRUE)

```


#Predicting Attrition
```{r}
employee_noattri = read.csv('CaseStudy2CompSet No Attrition.csv', header = TRUE) 

classifier1 = naiveBayes(model[,-c(1,14,24,28,5,21,23,9,10,11,37:45)],model$Attrition)


attrition_pred = predict(classifier1, employee_noattri)
employee_noattri$PredictAttrition = unlist(attrition_pred)


PredictionsAttritionDF = data.frame(c(employee_noattri$ID), c(employee_noattri$PredictAttrition))

#Checking Prediction with df
PredictionsAttritionDF %>% count(c.employee_noattri.PredictAttrition.)
employee_noattri %>% count(PredictAttrition)

#Plotting the Prediction
PredictionsAttritionDF %>% ggplot(aes(x = c.employee_noattri.PredictAttrition., fill=c.employee_noattri.PredictAttrition.)) + labs(title = "Predicting Attrition", x = "Predicted Attrition", fill = "Predicted Attrition") + geom_bar()

#Writing Prediction into CSV file 
write.csv(PredictionsAttritionDF, "AttritionPrediction.csv", row.names = TRUE)

```
